---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# dream: An R Package for Dynamic Relational Event Analysis and Modeling

<!-- badges: start -->
```{r, echo = FALSE, results='asis'}
library(badger)
cat(
	badge_cran_release("dream", "blue"),
	badge_cran_download("dream", "grand-total", "red"),
  badge_cran_checks("dream")
)
```

[![R-CMD-check](https://github.com/kevinCarson/dream/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/kevinCarson/dream/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The `dream` package provides users with helpful functions for relational event modeling/analysis. In particular, `dream` provides users with helper functions for large relational event analysis, such as recently proposed sampling procedures for creating relational risk sets. Alongside the set of functions for relational event analysis, this package includes functions for the structural analysis of one- and two-mode networks, such as network constraint and effective size measures. 

This package was developed with support from the National Science Foundation’s (NSF) Human Networks and Data Science Program (HNDS) under award number 2241536 (PI: Diego F. Leal). Any opinions, findings, and conclusions, or recommendations expressed in this material are those of the authors and do not necessarily reflect the views of the NSF.

## Authors

**Kevin A. Carson**  
  - Author & Maintainer    
  - PhD Candidate at the University of Arizona School of Sociology  
  - Email: kacarson@arizona.edu  
  - Website: https://kevincarson.github.io/  

**Diego F. Leal**  
  - Author & Maintainer  
  - Associate Professor at the University of Arizona School of Sociology  
  - Email: dflc@arizona.edu  
  - Website: https://www.diegoleal.info/index.html  

## Installation

You can install the stable verison of `dream` from [CRAN](https://cran.r-project.org/package=dream) via:  
``` r
install.packages("dream")
```

You can install the development version of `dream` from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("kevinCarson/dream")
```

## Estimating an (Ordinal) Relational Event Model in `dream`

### Sampling from the Observed Events and Case-Control Sampling

This is a basic example which shows how to sample from the observed events
and employ the case-control sampling technique for large relational event models (see [Butts 2008](https://onlinelibrary.wiley.com/doi/abs/10.1111/j.1467-9531.2008.00203.x)) following [Lerner and Lomi (2020)](https://www.cambridge.org/core/journals/network-science/article/reliability-of-relational-event-model-estimates-under-sampling-how-to-fit-a-relational-event-model-to-360-million-dyadic-events/B4286F370CD3A1A4ED30DF5120F04897) and [Vu et al. (2015)](https://www.sciencedirect.com/science/article/abs/pii/S0378873315000477). Then based upon the post-processing event sequence, the example computes a set of standard network statistics for two-mode relational event models. Lastly, the examples estimates an ordinal timing relational event model. The event sequence included in this example is based a subset (i.e., the first 100,000 events) of the 2018 Wikipedia article-edit event sequence used in [Lerner and Lomi (2020)](https://www.cambridge.org/core/journals/network-science/article/reliability-of-relational-event-model-estimates-under-sampling-how-to-fit-a-relational-event-model-to-360-million-dyadic-events/B4286F370CD3A1A4ED30DF5120F04897).

```{r processevents}
library(dream)
data("WikiEvent2018.first100k")
WikiEvent2018.first100k$time <- as.numeric(WikiEvent2018.first100k$time)
### Creating the EventSet By Employing Case-Control Sampling With M = 10 and
### Sampling from the Observed Event Sequence with P = 0.01
EventSet <- createriskset(
  type = "two-mode",
  time = WikiEvent2018.first100k$time, # The Time Variable
  eventID = WikiEvent2018.first100k$eventID, # The Event Sequence Variable
  sender = WikiEvent2018.first100k$user, # The Sender Variable
  receiver = WikiEvent2018.first100k$article, # The Receiver Variable
  p_samplingobserved = 0.10, # The Probability of Selection
  n_controls = 10, # The Number of Controls to Sample from the Full Risk Set
  seed = 9999) # The Seed for Replication

post.processing.riskset <- EventSet[EventSet$sampled == 1,] #only those sampled events! 
```

### A Miniature Replication of Lerner and Lomi (2020)

```{r computestats}
# computing the inertia statistic with the exponential weights and a halflife
# value of 30 days
post.processing.riskset$repetition <- repetition(
   time = EventSet$time,
   sender = EventSet$sender,
   receiver = EventSet$receiver,
   sampled = EventSet$sampled,
   observed = EventSet$observed,
   halflife = 2.592e+09, 
   dyadic_weight = 0,
   Lerneretal_2013 = FALSE)

# computing the sender outdegree statistic with the exponential weights and a halflife
# value of 30 days
post.processing.riskset$sender.outdegree <- degreestats(
   formation = "sender-outdegree",
   time = EventSet$time,
   sender = EventSet$sender,
   receiver = EventSet$receiver,
   sampled = EventSet$sampled,
   observed = EventSet$observed,
   halflife = 2.592e+09, 
   dyadic_weight = 0,
   Lerneretal_2013 = FALSE)

# computing the receiver indegree statistic with the exponential weights and a halflife
# value of 30 days
post.processing.riskset$receiver.indegree <- degreestats(
   formation = "receiver-indegree",
   time = EventSet$time,
   sender = EventSet$sender,
   receiver = EventSet$receiver,
   sampled = EventSet$sampled,
   observed = EventSet$observed,
   halflife = 2.592e+09, 
   dyadic_weight = 0,
   Lerneretal_2013 = FALSE)

# computing the four-cycles statistic with the exponential weights and a halflife
# value of 30 days
post.processing.riskset$fourcycles <- fourcycles(
   time = EventSet$time,
   sender = EventSet$sender,
   receiver = EventSet$receiver,
   sampled = EventSet$sampled,
   observed = EventSet$observed,
   halflife = 2.592e+09, 
   dyadic_weight = 0,
   Lerneretal_2013 = FALSE)

# Estimating the ordinal relational event model! 
lerner.lomi.rem <- remlogit(observed ~ 
                            repetition +
                            sender.outdegree + 
                            receiver.indegree + 
                            receiver.indegree:sender.outdegree +
                            fourcycles,
                            event.cluster = post.processing.riskset$eventID,
                            newton.rhapson=FALSE,
                            data = post.processing.riskset)
```

```{r, echo=TRUE}
summary(lerner.lomi.rem)
```
